### 2019-04-19 Manuscript Tex file/Texfile/Table1.tex
Init Commit

\begin{center}
\begin{table}[]
\begin{tabular}{lllll}
Pair & Weight & Community & Title                                                                                         &                                                             \\
1    & 6.03                      & FuncAn    & Properties of principal component methods for functional and longitudinal data analysis       &                                                             \\
     &                           & VarSel    & Nonconcave penalized likelihood with a diverging number of parameters                         &                                                             \\
2    & 3.08                      & MulT      & Innovated higher criticism for detecting sparse signals in correlated noise                   &                                                             \\
     &                           & VarSel    & Regularized estimation of large covariance matrices                                           &                                                             \\
3    & 2.47                      & DimRed    & Contour projected DimRed                                                                      &                                                             \\
     &                           & VarSel    & Factor profiled sure independence screening                                                   &                                                             \\
4    & 2.15                      & VarSel    & Factor profiled sure independence screening                                                   &                                                             \\
     &                           & DimRed    & Sliced regression for DimRed                                                                  &                                                             \\
5    & 1.86                      & VarSel    & Covariance regularization by thresholding                                                     &                                                             \\
     &                           & MulT      & Adapting to unknown sparsity by controlling the false discovery rate                          &                                                             \\
6    & 1.77                      & VarSel    & Asymptotic properties of bridge estimators in sparse high-dimensional regression models       &                                                             \\
     &                           & Mixed     & "Marginal asymptotics for the $''$large P                                                     & small n$''$ paradigm: with applications to microarray data" \\
7    & 1.71                      & VarSel    & A majorization-minimization approach to VarSel using spike and slab priors                    &                                                             \\
     &                           & Mixed     & Empirical \{B\}ayes selection of wavelet thresholds                                           &                                                             \\
8    & 1.66                      & VarSel    & Spades and mixture models                                                                     &                                                             \\
     &                           & MulT      & Adapting to unknown sparsity by controlling the false discovery rate                          &                                                             \\
9    & 1.61                      & DimRed    & A constructive approach to the estimation of DimRed directions                                &                                                             \\
     &                           & VarSel    & Factor profiled sure independence screening                                                   &                                                             \\
10   & 1.54                      & VarSel    & A majorization-minimization approach to VarSel using spike and slab priors                    &                                                             \\
     &                           & Mixed     & Spike and slab VarSel: frequentist and \{B\}ayesian strategies                                &                                                             \\
11   & 1.53                      & VarSel    & VarSel in nonparametric additive models                                                       &                                                             \\
     &                           & Mixed     & Nonparametric estimation of an additive model with a link function                            &                                                             \\
12   & 1.39                      & Mixed     & Nonparametric inferences for additive models                                                  &                                                             \\
     &                           & VarSel    & Nonparametric independence screening in sparse ultra-high-dimensional additive models         &                                                             \\
13   & 1.29                      & Mixed     & Bayesian VarSel in structured high-dimensional covariate spaces with applications in genomics &                                                             \\
     &                           & VarSel    & Model selection and estimation in regression with grouped variables                           &                                                             \\
14   & 1.20                      & VarSel    & The sparsity and bias of the \{LASSO\} selection in high-dimensional linear regression        &                                                             \\
     &                           & Mixed     & Tests for high-dimensional covariance matrices                                                &                                                             \\
15   & 1.19                      & FuncAn    & Empirical dynamics for longitudinal data                                                      &                                                             \\
     &                           & Mixed     & VarSel in nonparametric varying-coefficient models for analysis of repeated measurements      &                                                            
\end{tabular}
\end{table}
\end{center}